When are Fuzzy Extractors Possible?

### 2014-11-10 trunk/info-bounds/FuzzyExtPossibility.tex
presentation for brown
small typo fix

We now present an outline of the proof~(proofs in \apref{sec:fuzz ext proof}):
\begin{itemize}
\item \propref{prop:dist fuzzy ent fuzz}: for all $W\in \mathcal{W}$, $\Hfuzz(W) = \omega (\log n)$. That is, $\forall z, \Hfuzz(V | Z=z) = \omega(\log n)$.
\item \propref{prop:dist uniform fuzz}: the distribution $V$ is uniform.
\item \lemref{lem:fuzz can't get key}: In expectation across $Z$, a large subset of keys that are not possible.  In more detail,
\begin{itemize}
\item Half the keys have at most $2^{n- \kappa}$ pre images in the metric space~(this is at most half the metric space).  Denote this set as $R_{small}$.  
\item Consider some $\key \in R_{small}$.  %Our goal is to show there are few values of $v \in V|Z$ that could produce this key.  
Consider the set of $V_{\key } = \{w | \rep(w, p) = \key \}$.  All points in $V | \sketch(V)$ are distance $t$ from a boundary of $V_\key$~(the functionality of $\rep$ guarantees that for the true $w$ all nearby points map to the same $\key$).  We show that most of $V_\key$ is near a boundary.  A result of Frankel and F{\"u}redi says that the boundary of a region is minimized by a ball containing the same number of points~\cite{frankl1981short}.  Hoeffding's inequality says that most of a ball lies near its boundary~\cite{hoeffding1963probability}.  Together these two results imply that $V_\key$ is small.
\item As before, there are many possible values for $z_1, z_2$ for the side information $Z$~(and these possible values are equally likely).  Furthermore, the distributions $V|Z=z_1 $ and $V| Z=z_1$ have disjoint support outside of $v$.
\item For most values of possible $Z$, the intersection between the viable pre images of $V|Z$ and $V_\key$ contains at most one point~(the received point $v$).  Checking if $V|Z \cap V_{\key}$ is nonempty is an effective distinguisher.
%The previous two propositions were similar to the secure sketch counterexample.  However, the structure on the metric space imposed byThe remainder of the proof is significantly different.  We show for any $(\gen, \rep)$ many keys have few possible values that could produce them in the metric space once $p$ is published.  Then an adversary that sees $z$ is able to eliminate many possible keys and thus effectively distinguish the key from random.
\end{itemize}
\end{itemize}

@@

The optimum strategy is to include these values uniformly from different $Z$ values.
We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s= H_0(V | \sketch(V) = s)$.  %That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  
Recall that 
\[
\tilde{H}_0(V | \sketch(V)) =  \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \log \expe_{s\in \sketch(V)} 2^{h_s} %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as containers to be filled with the $2^{h_{ss}}$ items~(possible values of $V | \sketch(V)=ss$).  Each container receives automatically receives one free point~(all the distributions share $v$).  The average number of items in each container is maximized when the containers are filled equally.  That is, the average number of items in each container is bounded by the number of items divided by the number of container.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# items}+\text{\# containers}}{\text{\# containers}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1 \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following as long as $t\ge 4$~(using  \lemref{lem:log-minus-one}, which appears below):
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1\right)\\
&\le \max\left\{ \log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1, 1\right\}.
\end{align*}
Where the inequality follows because $\log x+1 \le \max\{ 1+ \log x,1\}$ for $x\ge 0$.
The left operand to $\max$ is bounded by $2$~(bounded the $\max$ by $2$):
\begin{align*}
\log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \left(\expe_{s\in \sketch(V)} 2^{h_s}\right) +1\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V)) +1 \\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+2\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+2\\
&< (\gamma-t+2)\log |\mathbb{F}| - (\gamma-2) \log |\mathbb{F}| +2 \ \ \ \ \mbox{(by \lemref{lem:log-minus-one})}\\
&\le (4-t)\log |\mathbb{F}| +2< 2\,.
\end{align*}
\end{proof}

>>>>

We now present an outline of the proof~(proofs in \apref{sec:fuzz ext proof}):
\begin{itemize}
\item \propref{prop:dist fuzzy ent fuzz}: for all $W\in \mathcal{W}$, $\Hfuzz(W) = \omega (\log n)$. That is, $\forall z, \Hfuzz(V | Z=z) = \omega(\log n)$.
\item \propref{prop:dist uniform fuzz}: the distribution $V$ is uniform.
\item \lemref{lem:fuzz can't get key}: In expectation across $Z$, a large subset of keys that are not possible.  In more detail,
\begin{itemize}
\item Half the keys have at most $2^{n- \kappa}$ pre images in the metric space~(this is at most half the metric space).  Denote this set as $R_{small}$.  
\item Consider some $\key \in R_{small}$.  %Our goal is to show there are few values of $v \in V|Z$ that could produce this key.  
Consider the set of $V_{\key } = \{w | \rep(w, p) = \key \}$.  All points in $V | \sketch(V)$ are distance $t$ from a boundary of $V_\key$~(the functionality of $\rep$ guarantees that for the true $w$ all nearby points map to the same $\key$).  We show that most of $V_\key$ is near a boundary.  A result of Frankel and F{\"u}redi says that the boundary of a region is minimized by a ball containing the same number of points~\cite{frankl1981short}.  Hoeffding's inequality says that most of a ball lies near its boundary~\cite{hoeffding1963probability}.  Together these two results imply that $V_\key$ is small.
\item As before, there are many possible values for $z_1, z_2$ for the side information $Z$~(and these possible values are equally likely).  Furthermore, the distributions $V|Z=z_1 $ and $V| Z=z_2$ have disjoint support outside of $v$.
\item For most values of possible $Z$, the intersection between the viable pre images of $V|Z$ and $V_\key$ contains at most one point~(the received point $v$).  Checking if $V|Z \cap V_{\key}$ is nonempty is an effective distinguisher.
%The previous two propositions were similar to the secure sketch counterexample.  However, the structure on the metric space imposed byThe remainder of the proof is significantly different.  We show for any $(\gen, \rep)$ many keys have few possible values that could produce them in the metric space once $p$ is published.  Then an adversary that sees $z$ is able to eliminate many possible keys and thus effectively distinguish the key from random.
\end{itemize}
\end{itemize}

@@

The optimum strategy is to include these values uniformly from different $Z$ values.
We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s= H_0(V | \sketch(V) = s)$.  %That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  
Recall that 
\[
\tilde{H}_0(V | \sketch(V)) =  \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \log \expe_{s\in \sketch(V)} 2^{h_s} %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as containers to be filled with the $2^{h_{ss}}$ items~(possible values of $V | \sketch(V)=ss$).  Each container receives automatically receives one free point~(all the distributions share $v$).  The average number of items in each container is maximized when the containers are filled equally.  That is, the average number of items in each container is bounded by the number of items divided by the number of container.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# items}+\text{\# containers}}{\text{\# containers}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1 \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following as long as $t\ge 4$~(using  \lemref{lem:log-minus-one}, which appears below):
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1\right)\\
&\le \max\left\{ \log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1, 1\right\}.
\end{align*}
Where the inequality follows because $\log x+1 \le \max\{ 1+ \log x,1\}$ for $x\ge 0$.
The left operand to $\max$ is bounded by $2$~(bounding the $\max$ by $2$):
\begin{align*}
\log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \left(\expe_{s\in \sketch(V)} 2^{h_s}\right) +1\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V)) +1 \\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+2\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+2\\
&< (\gamma-t+2)\log |\mathbb{F}| - (\gamma-2) \log |\mathbb{F}| +2 \ \ \ \ \mbox{(by \lemref{lem:log-minus-one})}\\
&\le (4-t)\log |\mathbb{F}| +2< 2\,.
\end{align*}
\end{proof}
